---
title: "Assessment of Social and Economical Impacts of Disasters in USA"
author: "Pranjal Mandhaniya"
date: "Thursday 22 January 2015"
output: html_document
---

### Synopsis

1. Unzipped archive and read CSV file under "wtrdt" name.
2. Examined dataframe for required names to analyze and subset it.
3. Check for fatalities and Injuries differently for disaster's impact on humans.
4. Similarly for economical consequences we saw crop damage and property damage differently.
5. Because of getting different results on both end we joined crop and property damage.

### Data Processing

1. Reading CSV file.

```{r,cache=TRUE}
wtrdt <- read.csv("repdata-data-StormData.csv") 
wtrdt <- wtrdt[,c("EVTYPE","FATALITIES","INJURIES","PROPDMG","CROPDMG")]
head(wtrdt)
summary(wtrdt)
```

2. Analyzing data for Human loss.

```{r, cache=TRUE }
ppldt <- aggregate(FATALITIES~EVTYPE,data = wtrdt,sum)
ppldt[ppldt$FATALITIES==max(ppldt$FATALITIES),]

njrdt <- aggregate(INJURIES~EVTYPE,data = wtrdt,sum)
njrdt[njrdt$INJURIES==max(njrdt$INJURIES),]

library(ggplot2)
ggplot(ppldt[ppldt$FATALITIES>500, ], aes(EVTYPE, FATALITIES)) + 
  geom_bar(stat = "identity", colour = "cyan",fill = "red", width = 1) + 
  labs(title = "Histogram of Fatalities v/s Events(Top)", x = "Event",y = "Total number of Fatalities") +
  theme(axis.text.x = element_text(angle = 90,hjust = 0.5,size = 10))
```

3. Analyzing for Economical loss.

```{r,cache=TRUE}
prpdt <- aggregate(PROPDMG~EVTYPE,data = wtrdt,sum)
prpdt[prpdt$PROPDMG==max(prpdt$PROPDMG),]

crpdt <- aggregate(CROPDMG~EVTYPE,data = wtrdt,sum)
crpdt[crpdt$CROPDMG==max(crpdt$CROPDMG),]

alldmgdt <- aggregate( CROPDMG+PROPDMG ~EVTYPE,data = wtrdt,sum)
c("EVTYPE","DAMAGE") -> names(alldmgdt) 
alldmgdt[alldmgdt[,2] == max(alldmgdt[,2]) , ]

ggplot(alldmgdt[alldmgdt[,2]>50000, ], aes(EVTYPE, DAMAGE)) + 
  geom_bar(stat = "identity", colour = "cyan",fill = "red", width = 1) + 
  labs(title = "Histogram of Damages v/s Events(Top)", x = "Event",y = "Total amount of Damages") +
  theme(axis.text.x = element_text(angle = 90,hjust = 0.5,size = 12))
```

### Results

1. Tornado cause major human loss and impact than any other event causing maximum fatalities and injuries.
2. Tornado also cause maximum property damage but Hail cause maximum crop damage.
3. As combination of crop and property damage its still Tornado on top.
